---
title: Session 5
output:
  ioslides_presentation:
    css: cust_css.css
  slidy_presentation: default
  beamer_presentation: default
subtitle: Data Visualization in R
author: "Tek Song"
ratio: 16x10
---
```{r setup, include=F}
knitr::opts_chunk$set(echo=F, eval=F, cache = T)
```

## Session outline
<ol>
<li>Base R plots</li>
<li>`ggplot2` package</li>
<li>Other useful packages</li>
</ol>

## Base R plots
<ol>
<li>Generic plot types in R</li>
<div class="row">
<div class="column">
<ul>
<li>Scatter plots</li>
<li>Histogram / density plots</li>
<li>Bar plots</li>
</ul>
</div>
<div class="column">
<ul>
<li>Box plots</li>
<li>Line plots</li>
<li>Pie charts</li>
</ul>
</div>
</div>
<!-- http://www.sthda.com/english/wiki/r-base-graphs -->
</ul>
<li>Combining multiple plots</li>
<ul>
<li>How to add a new layer on top of an existing plot</li>
<li>How to create a matrix of plots</li>
</ul>
<li>Saving graphs</li>
<ul>
<li>Format output</li>
</ul>
</ol>


## hrefs
<a href="https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/pairs">`pairs()`</a> function
<!-- <a href="https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/plot">`plot()`</a> function -->
<!-- <a href="https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/lines">`lines()`</a> function -->
<!-- <a href="https://www.rdocumentation.org/packages/alineR/versions/1.1.4/topics/aline">`aline()`</a> function -->
<!-- <a href="https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/hist">`hist()`</a> function -->
<!-- <a href="https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/boxplot">`boxplot()`</a> function -->
<!-- <a href="https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/barplot">`barplot()`</a> function -->
<!-- <a href="https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/pie">`pie()`</a> function -->
<!-- <a href="https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/density">`density()`</a> function -->
text

## args
main -- title of the plot
xlab / ylab
xlim / ylim
pch
frame
col  hex
break
horizontal


## layout
layout()
par() mfrow


## ggplot2
Template
```{r}
ggplot(data=<DATA>) + 
  <geom_function>(mapping=aes(<mappings>))
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))

# vs.
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```



```{r}
ggplot(data = diamonds) + 
  stat_count(mapping = aes(x = cut))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))

diamonds %>%
  group_by(cut) %>%
  summarize(freq=n()) %>%
  ggplot() +
  geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")
```

```{r}
df1<- read.csv("data/thermostat_amazon.csv")
scoring<- read.csv("data/Non_Smart_Thermostat.csv")
library(dplyr)
library(ggplot2)
df1$id=1:nrow(df1)
scoring %>%
  group_by(X1, topic) %>%
  summarise(avg_score=mean(sentiment2)) %>%
  tidyr::spread(key="topic", value="avg_score") %>%
  # scale() %>%
  select(X1, Costs, `Easy/Hard`, Feelings, Installation, NO)->score2
score2[score2==0]<-NA
left_join(df1, score2, by=c("id"="X1"))->df
# apply(score2, 2, function(x){sum(is.na(x))})

df %>%
  ggplot()+
  # geom_histogram(aes(x=Costs, fill=product), position = "stack")+
  # geom_histogram(aes(x=Costs, fill=product), position = "identity")+
  geom_histogram(aes(x=NO, fill=product), position = "dodge")+
  scale_y_continuous(labels = scales::percent)

df %>%
  ggplot()+
  geom_point(aes(x=`Costs`, y=`NO`))

mpg %>%
  group_by(class) %>%
  summarize(count=n())

# Bar Chart
mpg %>%
  count(class, name="count") %>%
  ggplot()+
  # geom_bar(aes(x=class, y=count), stat="identity")
  geom_col(aes(x=class, y=count))

mpg %>%
  ggplot()+
  stat_count(aes(x=class))

mpg %>%
  ggplot()+
  geom_bar(aes(x=class))

mpg %>%
  ggplot()+
  geom_bar(aes(
    x=class,
    y=after_stat(count/sum(count))*100    # Use after_stat() to modify mapping from stats
    ))

mpg %>%
  ggplot()+
  geom_density(
    aes(
      x=hwy,
      # y=after_stat(scaled)
      )
  )
??geom_density


mpg %>%
  ggplot(aes(x=class, y=hwy))+
  geom_boxplot()+
  stat_summary(fun=mean, geom="point", color="red")
# Based on the vector type for color argument, discrete or continuous color scale is picked


##### Scale
mpg %>%
  ggplot()+
  geom_point(aes(x=displ, y=hwy, color=class))+
  scale_color_brewer(type="qual")+  # color palettes   RColorBrewer::display.brewer.all()
  scale_x_continuous(breaks=c(3,5,6))+
  scale_y_continuous(trans="log10")     # x and y are also controlled with scales


mpg %>%
  ggplot()+
  geom_point(aes(x=displ, y=hwy, color=class, size=cyl))+
  scale_color_brewer(type="qual")+
  scale_size(breaks=c(4,5,6,8))


##### Facet
mpg %>%
  ggplot()+
  geom_point(aes(x=displ, y=hwy))+
  # facet_wrap(~class)                     # Facet is often the best way to avoid overplotting
  # facet_wrap(~ drv, scales='free')       # free_y or free_x
  # facet_grid(year ~ drv)                 # facet_grid graphic pivots
  facet_wrap(~ drv+year)                 # Subsetting data

mpg %>%
  ggplot()+
  geom_bar(aes(y=manufacturer))+
  facet_grid(class ~ ., scale="free_y" , space="free_y")



##### Coordinates
mpg %>%
  ggplot()+
  geom_bar(aes(x=class))+
  # coord_polar()
  # scale_y_continuous(limits=c(0, 40))      # compact, midsize, suv, Removed...
  coord_cartesian(ylim=c(0, 40))             # Zoom-in


mpg %>%
  ggplot()+
  geom_point(aes(x=hwy, y=displ))+
  # scale_y_continuous(trans="log")
  coord_trans(y="log")

# cartography?
# w <- sf::st_as_sf(maps::map('world'), plot=F, fill=T)  
# w %>%
#   ggplot()+
#   geom_sf()+
#   coord_sf(crs="")

### Theme
# There are built-in themes
mpg %>%
  ggplot()+
  geom_bar(aes(y=class))+
  facet_wrap(~year)+
  theme_minimal()+
  labs(
    title="this is a title",
    caption="this is a caption",
    x="this is x-axis label",
    y="this is y-axis label"
       )+
  theme(
    # text=element_text("times new roman"),                  # https://stackoverflow.com/questions/34522732/changing-fonts-in-ggplot2
    axis.text = element_text(colour = "red", size = rel(1.5), family = "serif"),
    panel.grid.major = element_line("steelblue"),
    # panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank()
  )
```


## ggplot2 topics
https://r4ds.had.co.nz/data-visualisation.html
facet_wrap / facet_grid