---
title: Session 2
output:
  ioslides_presentation:
    css: cust_css.css
  slidy_presentation: default
  beamer_presentation: default
subtitle: Data Type and Structure
author: "Tek Song"
ratio: 16x10
---
```{r setup, include=F}
knitr::opts_chunk$set(echo=T, eval=F, cache = T)
```

## Session Outline

<ol type="1">
<li>**Explore Data Types**</li>
-- Logical, Integer, Double, Character, <p style="color:LightGray; display:inline">Factor</p>, <p style="color:LightGray; display:inline">Date</p>
<ol>
<ol type="i">
<li>How to examine data type</li>
<li>How to assign data type</li>
</ol>
</ol>

<li>**Explore Data Structures**</li>
-- Vector, List, Matrix, Data Frame
<ol>
<ol type="i">
<li>How to create each data structure</li>
<li>How to modify each data structure</li>
<li>How to convert one data structure to another</li>
<li>How to explore/examine a given data</li>
</ol>
</ol>
</ol>


## Data Frame | How to create a data frame
<ol>
<li>Use vectors as building blocks -- <a href="https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/data.frame">`data.frame()` function</a></li>
```{r eval=T, results='hide'}
# Data frame is Heterogeneous!
MKT_dept <- data.frame(
  ID=1:7,
  Name=c("Felipe", "Guangzhi", "Minzhe", "SooYon", "Tek", "Yvonne", "Xiang"),
  Year=c(4, 1, 5, 2, NA, 1, 5),
  RA=factor(c("CB","Qt", "CB","CB", "Qt", "CB", "CB"), levels=c("CB", "Qt"))
) # str(MKT_dept)
```

```{r eval=T, echo=F, comment=""}
str(MKT_dept)
```
</ol>

## Data Frame | How to create a data frame
<ol start="2">
<li>Import data as data.frame</li>
<p style="font-size:18px">Data frame is the <em>de facto</em> data structure for most rectangular data set</p>
<ol type="i">
<li>`.csv()` extension</li>
```{r eval=T, comment=""}
df_csv <- read.csv("data/csv_example.csv")# csv example from Session 1
str(df_csv)
```

```{r}
readr::read_csv("data/csv_example.csv")  # Class: ... "data.frame" 
data.table::fread("data/csv_example.csv")# Class: ... "data.frame"
```
</ol>
</ol>


## Data Frame | How to create a data frame
<ol start="2">
<li>Import data as data.frame</li>
<ol type="i" start="2">
<li>`.xlsx()` extension</li>
```{r}
openxlsx::read.xlsx("data/xlsx_example.xlsx", 1)# class: "data.frame"
readxl::read_excel("data/xlsx_example.xlsx", 2) # class:..."data.frame"
```
</ol>
</ol>


## Data Frame | How to create a data frame  
<ol start="3">
<li>Convert a matrix to a data.frame -- <a href="https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/data.frame">`as.data.frame()`</a></li>
```{r eval=T, comment=""}
mdat=matrix(1:9, nrow=3, 
            dimnames = list(paste0("r", 1:3), paste0("c", 1:3)))
df <- as.data.frame(mdat)
c("is.data.frame" = is.data.frame(df),
  "is.list"       = is.list(df),
  "is.matrix"     = is.matrix(df),
  "is.atomic"     = is.atomic(df))
```
</ol>

## Data Frame | How to create a data frame  
<ol start="4">
<li>Convert a list to a data.frame</li>
<ol type="i">
<ol>Column Bind</ol>
```{r eval=T, comment=""}
a_list<-list(a=1:3, b=c("a", "b", "c"), c=factor(c(1,1,2)))
data.frame(a_list)
```

<ol>Row Bind</ol>
```{r}
do.call("rbind", a_list)


rng=function(x){
 sample(1:10, 5)
}
set.seed(1)
l<-lapply(1:5, rng)
data.table::rbindlist(l)

a_list<-list(1:3, c("a", "b", "c"), factor(c(1,1,2)))
data.table::rbindlist(a_list, fill=T)
# data.frame(Reduce("rbind", a_list))
# do.call("rbind", a_list)
```
</ol>